---
title: "Real estate analysis"
author: "Hatim Hussaini"
date: "2024-11-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploring Real Estate Data

### Understanding Relationships Between Variables

```{r load_dataset, include=FALSE}
real_estate <- read.csv("/Users/hatimhussaini/Downloads/Assignment datasets-20241012/realestate2024.csv")
```

```{r correlation_plot, echo=TRUE}
pairs(real_estate, main = "Scatter Plot Matrix of Real Estate Data", panel = panel.smooth)
```

***Insights:***

- Price vs. size has the strongest positive correlation, as bigger properties tend to be more expensive.
- Slight positive correlations exist between price and both bedrooms and age, suggesting newer homes or homes with more bedrooms are valued higher.

```{r corr_matrix}
cor(real_estate)
```

***Insights:***

- Strong correlation between price and size (0.7799), indicating size is a key factor influencing price.
- Correlations with other predictors, like bedrooms and age, are minimal.

### Building a Full Regression Model to Predict Property Price

#### Estimating the Impact of Property Size on Price

```{r fit-full-model, echo=TRUE}
model <- lm(price ~ size + bedrooms + age, data = real_estate)
summary(model)
summary_model <- summary(model)
```

$$
\hat{\beta}_{size} \pm t_{n - p, 1-\alpha/2} \cdot s.e.(\hat{\beta}_{size})
$$

$$
=\hat{\beta}_{size} \pm t_{193, 0.975} \cdot s.e.(\hat{\beta}_{size})
$$
$$
= 4.9371 \pm 1.97 \times 0.2819
$$
$$
=(4.38,5.49)
$$
We are 95% confident that for every unit increase in size, the price of the property will increase by between 4.38 and 5.49 thousand dollars on average.

### Testing the Overall Significance of the Model

#### Conducting an F-Test

**Theoretical Model:**

$$
Y_i = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3 + \epsilon_i, \quad i = 1, 2, \dots, n
$$

- $Y_i$ is the response variable (price).
- $\beta_0$ is the intercept.
- $\beta_1, \beta_2, \beta_3$ are coefficients for the predictors $X_1$ (size), $X_2$ (bedrooms), and $X_3$ (age).
- $\epsilon_i \sim N(0,\sigma^2)$ represents random variation.

**Null Hypothesis:** $H_0: \beta_1 = \beta_2 = \beta_3 = 0$ (no relationship between predictors and price).

```{r ANOVA}
anova(model)
```

- Regression SS = 50287541.
- Regression MS = 16762514.
- F-statistic: 105.43323.
- p-value: 1.9e-40 < 0.05.

***Conclusion:***
- There is significant evidence to reject $H_0$.
- At least one predictor (size, bedrooms, age) significantly impacts price.

### Validating the Regression Model

```{r Validation}
par(mfrow=c(1,2))  
plot(model, which=1:2)

plot(resid(model) ~ real_estate$size, xlab = 'Size', ylab = 'Residuals')
abline(h=0, col="red")
plot(resid(model) ~ real_estate$bedrooms, xlab = 'Bedroom', ylab = 'Residuals')
abline(h=0, col="red")
plot(resid(model) ~real_estate$age, xlab = 'Age', ylab = 'Residuals')
abline(h=0, col="red")
```

***Insights:***

- Residual vs. fitted plots show mostly random scatter with some imbalance in homoscedasticity.
- Q-Q plot shows deviations from normality.
- Residuals vs. predictors indicate linearity.

Overall, the model meets assumptions for linear regression with minor deviations.

### Evaluating Model Fit (R²)

$$
R^2 = 1 - \frac{SS_{Res}}{SS_{Total}} = 1 - \frac{30684511}{50287541} = 0.39
$$

An $R^2$ of 0.39 suggests the model explains 39% of the variability in property prices, indicating potential missing predictors like location.

### Refining the Model with Feature Selection

```{r best regression model 1}
summary_model <- summary(model)
summary_model
```

Removing the least significant predictor (age):

```{r best regression model 2}
model2 <- update(model, . ~ . - age)
summary_m2 <- summary(model2)
summary_m2
```

**Final Model Equation:**

$$
\hat{Y} = 459.8715 + 53.7265 \text{Bedrooms} + 4.9318 \text{Size}
$$

### Comparing $R^2$ and Adjusted $R^2$

```{r r^2 comparision}
print(summary_model$r.squared)
print(summary_m2$r.squared)
```

```{r adj r^2 comparison}
print(summary_model$adj.r.squared)
print(summary_m2$adj.r.squared)
```

- $R^2$ decreases slightly, showing reduced overall accuracy.
- Adjusted $R^2$ increases, indicating the model is now more efficient.
